Missing The Point

### 2020-10-15 Manuscript.Rmd
methods section manuscript (before updating r and rstudio)

```{r setup, include=FALSE}
# environment
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mice)
load("Data/results.Rdata")

# sim results plotting function
generic_plot <- function(v, name, ...){
  results %>%
    mutate(
      p = factor(p, levels = c(0.05, 0.5, 0.95), labels = c("5%", "50%", "95%")),
      mech = factor(mech, levels = c("MCAR", "MAR", "MNAR"))
      ) %>% 
  ggplot(aes(
    x=.data[["it"]], 
    y=.data[[v]], 
    color = .data[["p"]], 
    shape = .data[["mech"]], 
    linetype = .data[["mech"]]
    )) +
  geom_line(na.rm = TRUE) +
  labs(
    x = "Number of iterations",
    y = name,
    color = "Incomplete \n cases", 
    shape = "Missingness \n mechanism", 
    linetype = "Missingness \n mechanism") +
  theme_classic() + 
  scale_color_manual(values = c(
    #"#009E73", #green
    "#56B4E9", #blue
    #"#E69F00", #orange
    "#F0E442", #yellow
    "#D55E00"  #red
  ))
  # scale_color_manual(values = c(
  #   # '#228833', #green
  #   # '#66CCEE', #blue
  #   '#CCBB44', #yellow
  #   '#EE6677',  #red~ish
  #   '#AA3377'  #maroon
  #   ))
}
```

@@

We propose a novel multivariate parameter to check for non-convergence in iterative algorithms. Our aim is to show which method (parameter + diagnostic + threshold) is the most informative about non-convergence in the iterative imputation algorithm implemented in `mice`.

- Missing data has to be accommodated, usually filled in

- Filling in is algorithmic, which relies on convergence

- Diagnosing convergence is difficult, both visually and diagnostically

- In this paper we explore diagnostic methods for MICE, and show which method (parameter + diagnostic + threshold) is the most informative about convergence

@@

- What does non-convergence look like? What are the consequences? Refer to van Buuren (2018) instead of reproducing it.

-

@@

- 3x3 missingness conditions: MCAR, MAR, MNAR + 5, 50, 95% of cases incomplete

- outcomes: estimated regression coefficient, CIW, coverage, R2 --> multivariate is key, because univariate convergence does not guarantee multivariate convergence

- convergence parameters: chain means, chain variances, scientific parameter, lambda

- diagnostics: rhat, ac

@@

Magnitude of the bias in r^2 perfectly follows the missingness mechanisms: least bias for MCAR, most for MNAR. Worst effect of non-convergence for p=0.95, MNAR, but stable from it=5.

@@

# References

>>>>

```{r setup, include=FALSE}
# environment
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mice)
load("Data/results_25_75.Rdata")

# sim results plotting function
generic_plot <- function(v, name, ...){
  results %>%
    mutate(
      p = factor(p, levels = c(0.25, 0.5, 0.75), labels = c("25%", "50%", "75%")),
      mech = factor(mech, levels = c("MCAR", "MAR", "MNAR"))
      ) %>% 
  ggplot(aes(
    x=.data[["it"]], 
    y=.data[[v]], 
    color = .data[["p"]], 
    shape = .data[["mech"]], 
    linetype = .data[["mech"]]
    )) +
  geom_line(na.rm = TRUE) +
  labs(
    x = "Number of iterations",
    y = name,
    color = "Incomplete \n cases", 
    shape = "Missingness \n mechanism", 
    linetype = "Missingness \n mechanism") +
  theme_classic() + 
  scale_color_manual(values = c(
    #"#009E73", #green
    "#56B4E9", #blue
    #"#E69F00", #orange
    "#F0E442", #yellow
    "#D55E00"  #red
  ))
  # scale_color_manual(values = c(
  #   # '#228833', #green
  #   # '#66CCEE', #blue
  #   '#CCBB44', #yellow
  #   '#EE6677',  #red~ish
  #   '#AA3377'  #maroon
  #   ))
}
```

@@

We propose a novel multivariate parameter to check for non-convergence in iterative algorithms. Our aim is to show which method is the most informative about non-convergence in iterative imputation. [explain that method = parameter + diagnostic + interpretation]

<!-- - Missing data has to be accommodated, usually filled in -->

<!-- - Filling in is algorithmic, which relies on convergence -->

<!-- - Diagnosing convergence is difficult, both visually and diagnostically -->

<!-- - In this paper we explore diagnostic methods for MICE, and show which method (parameter + diagnostic + threshold) is the most informative about convergence  -->

@@

- Iterative imputation is a sort of MCMC algorithm, so it makes sense to investigate non-convergence in a similar fashion to typical MCMC algorithms.

- What does non-convergence look like? --> non-stat + non-mix in a certain parameter, e.g., chain means.

- What are the consequences? --> bias, under-coverage --> Refer to van Buuren (2018) instead of reproducing this

- Existing non-convergence identifiers for MCMC algorithms are not all applicable, e.g. N_eff is estimated using varcov matrix of the set of samples, while this is not the case in it im per se.

- We consider autocorr and rhat, and monitor four parameters: chain means, chain variances, the scientific estimand, and lambda.

@@

We investigate non-convergence of the MICE algorithm through model-based simulation in R (version 4.0.2; R Core Team 2020). The number of simulation repetitions is 1000, and the simulation set-up is summarized in the pseudo-code below

```
# simulation set-up (pseudo-code)
for each simulation repetition (1:1000)
  1. simulate complete data
  for each missingness condition (1:9)
    2. create missingness
    for each iteration (1:50)
     3. impute missingness
     4. perform analysis of scientific interest
     5. compute non-convergence diagnostics 
     6. pool results across imputations
     7. compute performance measures
 8. combine outcomes of all missingness conditions
9. aggregate outcomes of all simulation repetitions
```

### Aims
Our aim is to assess the impact of non-convergence on the validity of scientific estimates obtained using `mice`. To induce non-convergence we terminate the iterative algorithm at different imputation chain lengths.

### Data generating mechanism
Data are generated in each simulation repetition for a sample of $n_{obs} = 1000$ cases. Let the complete set be three multivariate normally distributed random variables

$$
\begin{pmatrix}
Y\\
X_1\\
X_2
\end{pmatrix}
\sim \mathcal{N}
\begin{bmatrix}
\begin{pmatrix}
0\\
0\\
0
\end{pmatrix},
\begin{pmatrix}
2     &       &  \\
0.5   & 1     &  \\
-0.5  & 0.5   & 1
\end{pmatrix}
\end{bmatrix}.
$$
The complete set is amputed according to a $3\times3$ factorial design, consisting of three missingness mechanisms and three proportions of incomplete cases. [Briefly describe MCAR, MAR, MNAR + 25, 50, 75% of cases incomplete (not using 5% anymore, because according to Vink, n.d., 25% is more representative of missing data problems in the social sciences)]

### Estimands
We perform multiple linear regression to predict outcome variable $Y$: $Y' = \beta_0 + \beta_1 X_1 + \beta_2 X_2$, where $Y'$ is the expected value of the outcome. Our estimands are regression coefficient $\beta_1$, and the coefficient of determination, $r^2$.

### Methods
We use a combination of two non-convergence identifiers---autocorrelation and $\widehat{R}$---and four parameters of interest---chain means, chain variances, a scientific estimate, and the novel parameter we propose, $\lambda$.

### Performance measures
As recommended by @buur18, our performance measures are bias, empirical coverage rate and confidence interval width of the estimands.

<!-- - outcomes: estimated regression coefficient, CIW, coverage, r2 multivariate is key, because univariate convergence does not guarantee multivariate convergence -->

<!-- - convergence parameters: chain means, chain variances, scientific parameter, lambda -->

<!-- - diagnostics: rhat, ac -->

@@

Magnitude of the bias in $r^2$ perfectly follows the missingness mechanisms: least bias for MCAR, most for MNAR. Worst effect of non-convergence for $p=0.95$, MNAR, but stable from it=5.

@@

# References (incomplete)
